Online {\IterComp}\\Guided by Low Overhead Profiling

### 2017-05-08 cgo18mscr.tex
Restructuring of the section Related Work; Extract discussion of previous work from other sections and place them in as subsections of Related Work;

\begin{document}
%\title{A work-based metric for efficiently applying online iterative optimisation}
\title{Online Iterative Optimisation\\Guided by Low Overhead Profiling}
%\titlenote{Produces the permission block, and
%  copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
%  \texttt{acmart.pdf} document}

@@

\begin{abstract}
Efficiently selecting the best combination of compiler optimisations for a program
remains an open problem due to their complex interaction and the large optimisation search space.
Iterative optimisation techniques, which consist of iterating
over this large space searching for the best optimisation,
has the ability to adapt to new platforms, programs and workload while still
having a systematic and simple optimisation process.
Although some recent work have studied the impact of using multiple
input datasets for performing iterative optimisation, applying this
technique in an online scenario still poses some challenges.

In this paper, our main goal is to enable iterative optimisation in
online scenarios with the restriction of executing distinct inputs
only once.
In order to address this challenge,
we propose a the use of a work-based metric that allows for comparing different
combination of compiler optimisations even when executed with distinct inputs.
Because of its online nature, we also propose a relaxed instrumentation
that performs low overhead profiling of the amount of work executed, while
being able to trade-off between accuracy and overhead in a controlled manner.
\end{abstract}

@@

\keywords{Iterative Optimisation, Profiling, Relaxed Instrumentation}

@@

A well known compilation technique that addresses this challenge is iterative
optimisation.
Iterative optimisation has the ability to adapt to new
platforms, program and workload while still
having a systematic and simple optimisation process.
Iterative optimisations repeatedly evaluate a large number
of compiler optimisations until the best combination is
found for a particular program~\cite{fursin07,chen10}.
The main challange concerning iterative optimisation is the need for quickly
exploring such a large optimisation space~\cite{fursin07,cavazos07,zhou12}.

Until recently, most of existing work  had been focusing
on finding the best optimisation through repeated runs using a single input.
Although they demonstrate the potential of iterative optimisation,
in real scenarios the user rarely executes the same input dataset multiple times~\cite{bodin98,kisuki99,stephenson03,kulkarni04,agakov06}.
Applying iterative optimisation in light of a single input may not result in good performance when executing the optimised code with different inputs.

Most of real world applications are complex enough 
so that a single input case does not capture the whole
range of possible scenarios and program behaviour~\cite{haneda06,fursin07,chen10,chen12a}.
Because programs can exhibit behaviours that differ greatly depending on the input,
%This is something that is well known when writing test
%cases for correctness.
%We can extend the same argument when optimising the code for performance.
using a single input for iterative optimisation can produce poor performance when executed with different inputs.

Recent work have been studying
the impact of using multiple input datasets
for performing iterative optimisation.
This previous work suggests that optimising based on a representative range of input datasets
allows for selecting a robust compiler optimisation 
that produces good performance in real scenarios where
the input is expected to change constantly~\cite{haneda06,fursin07,chen10,chen12a,chen12b,fang15,mpeis16}.
Their results show that a limited number of input datasets may be sufficient to
obtain a well optimised program for a wider range of different inputs~\cite{haneda06,fursin07,chen10,chen12a}.

Finding such a robust combination of compiler optimisations by means of iterative optimisation across a large range of inputs
may be very time consuming.
In order to speed up this process we intend to reduce the number of actual executions during the exploration
of the optimisation space without much degradation of the final selected optimisation.

\begin{figure}[htb]
    \centering
    \includegraphics[width=\linewidth]{figs/diagram}
    \caption{Overview of the execution engine for applying iterative optimisation.}
    \label{fig:diagram}
\end{figure}

In this paper, our main goal is to enable iterative optimisation in
\textit{online} scenarios with the restriction of executing distinct inputs
only once, while targeting optimal performance across different inputs.
This online aspect is usually found in mobile and data centre
platforms~\cite{chen12b,fang15,mpeis16}, where the goal is to optimise
programs based on the workload of a particular user (or group of users)
between executions.

@@

Our main contributions are the following:
\begin{itemize}
\item The use of a work-based metric in order to enable online iterative
optimisation by comparing different combination of compiler optimisations even
when executed with distinct inputs.
\item We propose a relaxed instrumentation for low overhead profiling, with a controlled
trade-off between accuracy and overhead.
\end{itemize}

Chen~\etal~\cite{chen10,chen12a} evaluate the effectiveness of iterative
compilations across a large number of input test cases.
Their main motivation is to answer the question:
\textit{How data input dependent is iterative optimisation?}
Their results show that it is possible to find a combination of compiler optimisations
that achieves at least 86\% of the maximum speedup across all input test cases.
These optimal combinations are program-specific and yield average
speedups up to 3.75$\times$ over the highest optimisation level of compilers.

When optimising a program, the main method for iterative optimisation used by Chen~\etal~\cite{chen10,chen12a}
evaluates each combination of compiler optimisations across all the available inputs, i.e.,
if $N$ is the number of input test cases and $M$ is the total number of combinations of compiler optimisations,
they perform a total of $O(NM)$ runs of the program being optimised.
Furthermore, they use a pre-defined set of only 300 different combinations of compiler optimisations,
which represents a very small sample of the optimisation search space for most modern compilers, e.g.
LLVM has 56 distinct optimisation passes and GCC has about 47 high-level (SSA form) optimisation passes and
about 25 low-level (RTL) optimisation passes, which in both cases result in much more than $2^{50}$ distinct
combinations of compiler optimisations, without considering repetition.

Recent work~\cite{chen12b,fang15} have applied the same idea of performing
input-dependent iterative optimisation to distributed applications on data centres.
In summary, each worker receives a subset of the input dataset, called the evaluation dataset,
to perform an \textit{online} iterative optimisation of the code being executed.
Each worker performs the same the method for iterative optimisation used by Chen~\etal~\cite{chen10,chen12a},
i.e., they evaluate each combination of compiler optimisations across all the evaluation dataset.
However, because the optimisation is performed online, they usually consider a small evaluation dataset
and a small number of compiler optimisations.

Fursin~\etal~\cite{fursin07} addressed the problem of comparing the effect of two optimisations
on two distinct inputs. For that purpose, they proposed to use instruction per cycle (IPC) as the metric for performing such comparison.
Their result show that using IPC seems promising as a robust metric for iterative optimisation
across large input datasets.
However, 
some specific optimisation techniques may affect the use of IPC as a robust metric, and specially
IPC has been shown to provide poor performance estimation for multi-threaded programs~\cite{alameldeen06,eyerman08}.
In particular, IPC can give a skewed performance measure if threads spend time in \textit{spin-lock loops}
or other synchronisation mechanisms. 
Some existing work on performance assessment suggest that
total execution time should be used for measuring performance of multi-threaded programs~\cite{alameldeen06,eyerman08}.
Aalameldeen and Wood~\cite{alameldeen06} in particular suggest that a simple work-related metric should be used
if the unit of work is representative enough.
Work-related metrics have already been largely used for measuring performance of throughput-oriented applications,
for other applications, however, choosing an appropriate unit of work can be more challanging~\cite{alameldeen06}.

@@

A possible candidate for a work-based metric, borrowing from Physics,
is using the concept of \textit{average power}, i.e. the amount of
\textit{work}, $\Delta W$, performed during a period of time, $\Delta t$.
\[
   P_{avg} = \frac{\Delta W}{\Delta t}
\]

%The hypothesis is that, given two optimisations $o_1$ and $o_2$,
%a program compiled with optimisation $o_2$ would \textit{consistently} perform better
%than when it is compiled with $o_1$, i.e., $o_1 \tilde{<}_p \, o_2$,
%if it performs more \textit{work} per unit of time when compiled with $o_2$
%instead of $o_1$. 
By measuring the amount of \textit{work} done per unit of time we reduce
the impact of input-dependent aspects and focus instead on the efficiency
of the optimised program.
For this metric, the main challange is to precisely define what
represents \textit{work}.

@@

%In addition to being useful for speeding up iterative optimisation across a large
%number of different inputs, this metric based on comparing work-based efficiencies could
%potentially be used for applying iterative optimisation \textit{online}. 
%However, the overhead for estimating the amount of \textit{work} must be low
%enough for the performance gains to be beneficial.

@@

@@

In this section we describe how the computation of the work-metric can be
performed during runtime by means of instrumenting the code.
%with the necessary computation for the work metric, $\Delta W$, as described in Section~\ref{sec:metric}.
Because we define work as a linear equation on the block frequency counters,
it is possible to embed its computation into the execution of the program.
A naive instrumentation would consists basically of having a global counter
that starts with the interception value, $\varepsilon$, and each basic block
increments its own cost into the global counter. Although this instrumentation
is easily implemented, it inserts a large amount of overhead into the program.

In 1973, Donald Knuth~\cite{knuth73} provided an optimal algorithm for
profiling block frequency, inserting fewer probes than the previously described
naive instrumentation that was commonly used in practice~\cite{knuth71}.
Forman~\cite{forman81} proposed further overhead reductions by placing the
probes in basic blocks that are less likely to be executed based on static
heuristics.
Ball and Larus~\cite{ball94} offered a detailed discussion comparing
two approaches for optimally profiling block frequency, namely,
placing probes on the vertices or the edges of the control-flow graph (CFG).
They show that the edge-based approach produces optimal placement of probes.

@@

In this section we discuss our experimental evaluation.
First we describe the benchmarks with datasets used in the experiments.
Afterwards, we discuss our results concerning the instrumentations for profiling
our work metric.
Finally we present the results of the online iterative optimisation.

@@

%Iterative optimisations is essentially based on repeatedly trying
%out a large number of compiler optimisations until the best
%combination of compiler optimisations is found for a particular program.
%Our main goal is to speed up iterative optimisation while targeting
%optimal performance across large input datasets.

@@

%\vspace{1ex}
%\noindent \textbf{Iterative optimisation over a single input}

%Most of the existing iterative optimisation studies find the
%best optimisation though repeated runs on the same input.
%Although this approach will usually lead to sub-optimal performance
%across large input datasets, it provides a good baseline when considering
%the complexity and resonably low compile-time regarding
%iterative optimising compilers.
%If $M$ is the total number of combinations of compiler optimisations,
%this approach requires $O(M)$ runs of the program being optimised.
%We can consider two main scenarios:
%\textit{($i$ - best case scenario)}
%after selecting the optimisation over each individual input,
%consider the one with best performance across the whole input dataset;
%\textit{($ii$ - expected scenario)}
%after selecting the optimisation over each individual input,
%consider the average case of their performance across the whole input dataset.

%\vspace{1ex}
%\noindent \textbf{Iterative optimisation across large input datasets}

%Recent work on iterative optimisation have been targeting optimisation
%across multiple inputs~\cite{fursin07,chen10,chen12a}.
%If $N$ is the number of input test cases and $M$ is the total number of combinations of compiler optimisations,
%they perform a total of $O(NM)$ runs of the program being optimised.
%Similarly to what have been discussed previously,
%there are different ways for how to determine the optimal
%combination of compiler optimisations across multiple inputs.
%It is possible to tune the selection of the program-optimal combination
%to minimise risk or to maximise average performance.
%A compromise-based selection criterion could be to maximise average speedup
%with minimised variance.

%\vspace{1ex}
%\noindent \textbf{Iterative optimisation based on the IPC metric}

%While the previous two baselines addresses two opposite aspects of iterative optimisation,
%namely, compile-time efficiency and performance of the generated code,
%we also intend to compare against IPC as the competing baseline.
%The main reason for comparing against IPC is because it has been proposed as a metric
%for comparing the performance of two optimisations running on two distinct inputs.
%If $M$ is the total number of combinations of compiler optimisations,
%this approach requires $O(M)$ runs of the program being optimised.

\subsection{Evaluation of the Online Iterative Optimisation}

\begin{itemize}
\item Oracle-RM executes the program twice, for each input, and then measuring
the real speedup for each compiler optimisation and uses the real speedup for
selecting the best optimisation.
\item Oracle-PP represents the \textit{perfect} non-intrusive profiling by also
executing the program twice, once for estimating the amount of work and the
second for measuring its execution time.
This version uses the work-based metric, $\frac{\Delta W}{\Delta t}$, during
the iterative optimisation search.
\end{itemize}

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\textwidth]{figs/speedups.pdf}
    \caption{Speedups observed with the online iterative optimisation.}
    \label{fig:speedups}
\end{figure*}

%\begin{figure*}[htb]
%    \centering
%    \includegraphics[width=\textwidth]{figs/profiled-speedups.pdf}
%    \caption{Speedups observed with the online iterative optimisation if we
%             consider the instrumentation overhead.}
%    \label{fig:profiled-speedups}
%\end{figure*}

%\subsubsection{Contribution of individual optimisation passes}
%
%Figure~\ref{fig:flagsfreq} shows an aggregated view of the final combination
%of compiler optimisations that were selected by iterative optimisation search.
%The figure presents the individual optimisation passes with at least 1\% of
%frequency in the selected combination of compiler optimisations that improved
%the performance over {\texttt{-O3}}.
%
%\begin{figure*}[htb]
%    \centering
%    \includegraphics[width=\textwidth]{figs/flagsfreq.pdf}
%    \caption{Frequency of individual optimisation passes on the final selected 
%             compiler optimisations of the iterative optimisation search over
%             all benchmarks.}
%    \label{fig:flagsfreq}
%\end{figure*}

>>>>

\newcommand{\itercomp}{{iterative compilation}}
\newcommand{\Itercomp}{{Iterative compilation}}
\newcommand{\IterComp}{{Iterative Compilation}}

\begin{document}
%\title{A work-based metric for efficiently applying online iterative optimisation}
\title{Online {\IterComp}\\Guided by Low Overhead Profiling}
%\titlenote{Produces the permission block, and
%  copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
%  \texttt{acmart.pdf} document}

@@

\begin{abstract}
Efficiently selecting the best combination of compiler optimisations for a program
remains an open problem due to their complex interaction and the large optimisation search space.
{\Itercomp} techniques, which consist of iterating
over this large space searching for the best optimisation,
has the ability to adapt to new platforms, programs and workload while still
having a systematic and simple optimisation process.
Although some recent work have studied the impact of using multiple
input datasets for performing \itercomp, applying this
technique in an online scenario still poses some challenges.

In this paper, our main goal is to enable \itercomp in
online scenarios with the restriction of executing distinct inputs
only once.
In order to address this challenge,
we propose a the use of a work-based metric that allows for comparing different
combination of compiler optimisations even when executed with distinct inputs.
Because of its online nature, we also propose a relaxed instrumentation
that performs low overhead profiling of the amount of work executed, while
being able to trade-off between accuracy and overhead in a controlled manner.
\end{abstract}

@@

\keywords{{\IterComp}, Profiling, Relaxed Instrumentation}

@@

A well known compilation technique that
addresses this challenge is {\itercomp}.
{\Itercomp} has the ability to adapt to new
platforms, program and workload while still
having a systematic and simple optimisation process.
{\Itercomp} repeatedly evaluates a large number
of compiler optimisations until the best combination is
found for a particular program~\cite{fursin07,chen10}.
The main challenge concerning {\itercomp} is the need for quickly
exploring such a large optimisation space~\cite{fursin07,cavazos07,zhou12}.

Until recently, most of existing work  had been focusing
on finding the best optimisation through repeated runs using a single input.
Although they demonstrate the potential of {\itercomp},
in real scenarios the user rarely executes the same input dataset multiple times~\cite{bodin98,kisuki99,stephenson03,kulkarni04,agakov06}.
Applying {\itercomp} in light of a single input may not result in good performance when executing the optimised code with different inputs.

Most of real world applications are complex enough 
so that a single input case does not capture the whole
range of possible scenarios and program behaviour~\cite{haneda06,fursin07,chen10,chen12a}.
Because programs can exhibit behaviours that differ greatly depending on the input,
%This is something that is well known when writing test
%cases for correctness.
%We can extend the same argument when optimising the code for performance.
using a single input for {\itercomp} can produce poor performance when executed with different inputs.

Recent work have been studying
the impact of using multiple input datasets
for performing {\itercomp}.
This previous work suggests that optimising based on a representative range of input datasets
allows for selecting a robust compiler optimisation 
that produces good performance in real scenarios where
the input is expected to change constantly~\cite{haneda06,fursin07,chen10,chen12a,chen12b,fang15,mpeis16}.
Their results show that a limited number of input datasets may be sufficient to
obtain a well optimised program for a wider range of different inputs~\cite{haneda06,fursin07,chen10,chen12a}.

Finding such a robust combination of compiler optimisations by means of {\itercomp} across a large range of inputs
may be very time consuming.
In order to speed up this process we intend to reduce the number of actual executions during the exploration
of the optimisation space without much degradation of the final selected optimisation.

\begin{figure}[htb]
    \centering
    \includegraphics[width=\linewidth]{figs/diagram}
    \caption{Overview of the execution engine for applying {\itercomp}.}
    \label{fig:diagram}
\end{figure}

In this paper, our main goal is to enable {\itercomp} in
\textit{online} scenarios with the restriction of executing distinct inputs
only once, while targeting optimal performance across different inputs.
This online aspect is usually found in mobile and data centre
platforms~\cite{chen12b,fang15,mpeis16}, where the goal is to optimise
programs based on the workload of a particular user (or group of users)
between executions.

@@

Our main contributions are the following:
\begin{itemize}
\item The use of a work-based performance metric in order to enable
\textit{online} {\itercomp} by comparing different combination of
compiler optimisations even when executed with distinct inputs.
\item We propose a relaxed instrumentation for low overhead profiling, with a controlled
trade-off between accuracy and overhead.
\end{itemize}

@@

In this section we define the work-based performance metric used in this paper.
We define the performance metric as the ratio between the amount of
\textit{work}, $\Delta W$, performed during a period of time, $\Delta t$.
\[
   P = \frac{\Delta W}{\Delta t}
\]

%The hypothesis is that, given two optimisations $o_1$ and $o_2$,
%a program compiled with optimisation $o_2$ would \textit{consistently} perform better
%than when it is compiled with $o_1$, i.e., $o_1 \tilde{<}_p \, o_2$,
%if it performs more \textit{work} per unit of time when compiled with $o_2$
%instead of $o_1$. 
By measuring the amount of \textit{work} done per unit of time we reduce
the impact of input-dependent aspects and focus instead on the efficiency
of the optimised program.
For this metric, the main challenge is to precisely define what
represents \textit{work}.

@@

%In addition to being useful for speeding up {\itercomp} across a large
%number of different inputs, this metric based on comparing work-based efficiencies could
%potentially be used for applying {\itercomp} \textit{online}. 
%However, the overhead for estimating the amount of \textit{work} must be low
%enough for the performance gains to be beneficial.

@@

Similarly to previous work~\cite{giusto01,powell09,brandolese11},
we derive the cost model for the instruction set by modeling the
problem as a multi-variable linear regression, where the
\textit{regression coefficients} are the costs of the instructions
and the \textit{regressors} (or \textit{explanatory variables})
are computed as $\sum_B N_B(i)f(B)$ for each instruction $i$.

@@

In this section we describe how the computation of the work-metric can be
performed during runtime by means of instrumenting the code.
In particular, we adapt the optimal algorithm proposed originally 
for profiling block frequency~\cite{knuth73,ball94}.
Afterwards, we propose a relaxed instrumentation that focus on
further reducing the overhead by considering the trade-off between
profiling accuracy and instrumentation overhead.

Because we define work as a linear equation on the block frequency counters,
it is possible to embed its computation into the execution of the program.
A naive instrumentation would consists basically of having a global counter
that starts with the interception value, $\varepsilon$, and each basic block
increments its own cost into the global counter. Although this instrumentation
is easily implemented, it inserts a large amount of overhead into the program.
However, it is possible to insert fewer probes by carefully placing the probes
in a way that is possible to reconstruct
the complete profiling information~\cite{knuth73,ball94}.

@@

In this section we discuss our experimental evaluation.
First we describe the benchmarks with datasets used in the experiments.
Afterwards, we discuss our results concerning the instrumentations for profiling
our work metric.
Finally we present the results of the online {\itercomp}.

@@

%{\Itercomp} is essentially based on repeatedly trying
%out a large number of compiler optimisations until the best
%combination of compiler optimisations is found for a particular program.
%Our main goal is to speed up {\itercomp} while targeting
%optimal performance across large input datasets.

@@

%\vspace{1ex}
%\noindent \textbf{{\Itercomp} over a single input}

%Most of the existing {\itercomp} studies find the
%best optimisation though repeated runs on the same input.
%Although this approach will usually lead to sub-optimal performance
%across large input datasets, it provides a good baseline when considering
%the complexity and resonably low compile-time regarding
%iterative optimising compilers.
%If $M$ is the total number of combinations of compiler optimisations,
%this approach requires $O(M)$ runs of the program being optimised.
%We can consider two main scenarios:
%\textit{($i$ - best case scenario)}
%after selecting the optimisation over each individual input,
%consider the one with best performance across the whole input dataset;
%\textit{($ii$ - expected scenario)}
%after selecting the optimisation over each individual input,
%consider the average case of their performance across the whole input dataset.

%\vspace{1ex}
%\noindent \textbf{{\Itercomp} across large input datasets}

%Recent work on {\itercomp} have been targeting optimisation
%across multiple inputs~\cite{fursin07,chen10,chen12a}.
%If $N$ is the number of input test cases and $M$ is the total number of combinations of compiler optimisations,
%they perform a total of $O(NM)$ runs of the program being optimised.
%Similarly to what have been discussed previously,
%there are different ways for how to determine the optimal
%combination of compiler optimisations across multiple inputs.
%It is possible to tune the selection of the program-optimal combination
%to minimise risk or to maximise average performance.
%A compromise-based selection criterion could be to maximise average speedup
%with minimised variance.

%\vspace{1ex}
%\noindent \textbf{{\Itercomp} based on the IPC metric}

%While the previous two baselines addresses two opposite aspects of {\itercomp},
%namely, compile-time efficiency and performance of the generated code,
%we also intend to compare against IPC as the competing baseline.
%The main reason for comparing against IPC is because it has been proposed as a metric
%for comparing the performance of two optimisations running on two distinct inputs.
%If $M$ is the total number of combinations of compiler optimisations,
%this approach requires $O(M)$ runs of the program being optimised.

\subsection{Evaluation of the Online {\IterComp}}

\begin{itemize}
\item Oracle-RM executes the program twice, for each input, and then measuring
the real speedup for each compiler optimisation and uses the real speedup for
selecting the best optimisation.
\item Oracle-PP represents the \textit{perfect} non-intrusive profiling by also
executing the program twice, once for estimating the amount of work and the
second for measuring its execution time.
This version uses the work-based metric, $\frac{\Delta W}{\Delta t}$, during
the {\itercomp} search.
\end{itemize}

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\textwidth]{figs/speedups.pdf}
    \caption{Speedups observed with the online {\itercomp}.}
    \label{fig:speedups}
\end{figure*}

%\begin{figure*}[htb]
%    \centering
%    \includegraphics[width=\textwidth]{figs/profiled-speedups.pdf}
%    \caption{Speedups observed with the online {\itercomp} if we
%             consider the instrumentation overhead.}
%    \label{fig:profiled-speedups}
%\end{figure*}

%\subsubsection{Contribution of individual optimisation passes}
%
%Figure~\ref{fig:flagsfreq} shows an aggregated view of the final combination
%of compiler optimisations that were selected by {\itercomp} search.
%The figure presents the individual optimisation passes with at least 1\% of
%frequency in the selected combination of compiler optimisations that improved
%the performance over {\texttt{-O3}}.
%
%\begin{figure*}[htb]
%    \centering
%    \includegraphics[width=\textwidth]{figs/flagsfreq.pdf}
%    \caption{Frequency of individual optimisation passes on the final selected 
%             compiler optimisations of the {\itercomp} search over
%             all benchmarks.}
%    \label{fig:flagsfreq}
%\end{figure*}

\subsection{{\IterComp}}

Chen~\etal~\cite{chen10,chen12a} evaluate the effectiveness of iterative
compilations across a large number of input test cases.
Their main motivation is to answer the question:
\textit{How data input dependent is {\itercomp}?}
Their results show that it is possible to find a combination of compiler optimisations
that achieves at least 86\% of the maximum speedup across all input test cases.
These optimal combinations are program-specific and yield average
speedups up to 3.75$\times$ over the highest optimisation level of compilers.

When optimising a program, the main method for {\itercomp} used by Chen~\etal~\cite{chen10,chen12a}
evaluates each combination of compiler optimisations across all the available inputs, i.e.,
if $N$ is the number of input test cases and $M$ is the total number of combinations of compiler optimisations,
they perform a total of $O(NM)$ runs of the program being optimised.
Furthermore, they use a pre-defined set of only 300 different combinations of compiler optimisations,
which represents a very small sample of the optimisation search space for most modern compilers, e.g.
LLVM has 56 distinct optimisation passes and GCC has about 47 high-level (SSA form) optimisation passes and
about 25 low-level (RTL) optimisation passes, which in both cases result in much more than $2^{50}$ distinct
combinations of compiler optimisations, without considering repetition.

Recent work~\cite{chen12b,fang15} have applied the same idea of performing
input-dependent {\itercomp} to distributed applications on data centres.
In summary, each worker receives a subset of the input dataset, called the evaluation dataset,
to perform an \textit{online} {\itercomp} of the code being executed.
Each worker performs the same the method for {\itercomp} used by Chen~\etal~\cite{chen10,chen12a},
i.e., they evaluate each combination of compiler optimisations across all the evaluation dataset.
However, because the optimisation is performed online, they usually consider a small evaluation dataset
and a small number of compiler optimisations.

Fursin~\etal~\cite{fursin07} addressed the problem of comparing the effect of two optimisations
on two distinct inputs. For that purpose, they proposed to use instruction per cycle (IPC) as the metric for performing such comparison.
Their result show that using IPC seems promising as a robust metric for {\itercomp}
across large input datasets.
However, 
some specific optimisation techniques may affect the use of IPC as a robust metric, and specially
IPC has been shown to provide poor performance estimation for multi-threaded programs~\cite{alameldeen06,eyerman08}.
In particular, IPC can give a skewed performance measure if threads spend time in \textit{spin-lock loops}
or other synchronisation mechanisms. 
Some existing work on performance assessment suggest that
total execution time should be used for measuring performance of multi-threaded programs~\cite{alameldeen06,eyerman08}.
Aalameldeen and Wood~\cite{alameldeen06} in particular suggest that a simple work-related metric should be used
if the unit of work is representative enough.
Work-related metrics have already been largely used for measuring performance of throughput-oriented applications,
for other applications, however, choosing an appropriate unit of work can be more challanging~\cite{alameldeen06}.

\subsection{Work and Input Size Metrics}

\subsection{Optimal Instrumentation}

In 1973, Donald Knuth~\cite{knuth73} provided an optimal algorithm for
profiling block frequency, inserting fewer probes than the previously described
naive instrumentation that was commonly used in practice~\cite{knuth71}.
Forman~\cite{forman81} proposed further overhead reductions by placing the
probes in basic blocks that are less likely to be executed based on static
heuristics.

Ball and Larus~\cite{ball94} offered a detailed discussion comparing
two approaches for optimally profiling block frequency, namely,
placing probes on the vertices or the edges of the control-flow graph (CFG).
They show that the edge-based approach produces optimal placement of probes.
